---
title: "Implementation of the K-Medoids (PAM) and CLARA Algorithms to Cluster Emissions from Gas Turbine"
author: "Kristian Ekachandra"
date: "2021"
output: openintro::lab_report
---

# Load All Library

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readxl)
library(Amelia)
library(knitr)
library(openintro)
library(dplyr)
library(ggplot2)
library(GGally)
library(caret) #confusion matrix
library(cluster) #cluster analysis
library(factoextra) #cluster visualization
library(clValid) #cluster validation
```

# Read Dataset and Exploratory Data Analysis

```{r, warning=FALSE, message=FALSE}
dat <- readxl::read_excel("dataset.xlsx")
summary(dat)

#Missing Value
Amelia::missmap(data.frame(dat)) #No Missing Value (100% Observed)

str(dat)
head(dat, n = 5)
tail(dat, n = 5)

#About the data:
##Size or total observation: 7384
##All variable is numerical

#Visualitation with Density Plot
##AT (Ambient Temperature)
ggplot2::ggplot(dat, aes(x = AT)) + 
  geom_density(fill = "lightblue") + 
  geom_vline(aes(xintercept = mean(AT)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Ambient Temperature (AT)")

##AP (Ambient Pressure)
ggplot2::ggplot(dat, aes(x = AP)) + 
  geom_density(fill = "lightgreen") + 
  geom_vline(aes(xintercept = mean(AP)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Ambient Pressure (AP)")

##AH (Ambient Humidity)
ggplot2::ggplot(dat, aes(x = AH)) + 
  geom_density(fill = "coral") + 
  geom_vline(aes(xintercept = mean(AH)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Ambient Humidity (AH)")

##AFDP (Air Filter Difference Pressure)
ggplot2::ggplot(dat, aes(x = AFDP)) + 
  geom_density(fill = "cyan") + 
  geom_vline(aes(xintercept = mean(AFDP)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Air Filter Difference Pressure (AFDP)")

##GTEP (Gas Turbine Exhaust Pressure)
ggplot2::ggplot(dat, aes(x = GTEP)) + 
  geom_density(fill = "red") + 
  geom_vline(aes(xintercept = mean(GTEP)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Gas Turbine Exhaust Pressure (GTEP)")

##TIT (Turbine Inlet Temperature)
ggplot2::ggplot(dat, aes(x = TIT)) + 
  geom_density(fill = "darksalmon") + 
  geom_vline(aes(xintercept = mean(TIT)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Turbine Inlet Temperature (TIT)")

##TAT (Turbine After Temperature)
ggplot2::ggplot(dat, aes(x = TAT)) + 
  geom_density(fill = "brown") + 
  geom_vline(aes(xintercept = mean(TAT)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Turbine After Temperature (TAT)")

##TEY (Turbine Energy Yield)
ggplot2::ggplot(dat, aes(x = TEY)) + 
  geom_density(fill = "azure") + 
  geom_vline(aes(xintercept = mean(TEY)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Turbine Energy Yield (TEY)")

##CDP (Compressor Discharge Pressure)
ggplot2::ggplot(dat, aes(CDP)) + 
  geom_density(fill = "violet") + 
  geom_vline(aes(xintercept = mean(CDP)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Compressor Discharge Pressure (CDP)")

##CO (Carbon Monoxide)
ggplot2::ggplot(dat, aes(x = CO)) + 
  geom_density(fill = "pink") + 
  geom_vline(aes(xintercept = mean(CO)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Carbon Monoxide (CO)")

##NOX (Nitrogen Oxides)
ggplot2::ggplot(dat, aes(x = NOX)) + 
  geom_density(fill = "orange") + 
  geom_vline(aes(xintercept = mean(NOX)), color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Nitrogen Oxides (NOX)")

#Correlation Plot
plot(dat, col = dat$NOX)
GGally::ggpairs(data.frame(dat))

#Scaling
dat <- scale(dat, center = TRUE, scale = TRUE)
dat %>% head()
```

# Split Data

```{r, warning=FALSE, message=FALSE}
set.seed(43961)
samp <- sample(nrow(dat), 0.7 * nrow(dat), replace = FALSE)
data_train <- dat[samp, ] #70% data_tain
data_test <- dat[-samp, ] #30% data_test
```

#Algorithms

## 1. K-Medoids (PAM)

```{r, warning=FALSE, message=FALSE}
# Choose optimum number of cluster
set.seed(43961)
factoextra::fviz_nbclust(data_train, FUNcluster = pam, method = "silhouette") #k = 2
factoextra::fviz_nbclust(data_train, FUNcluster = pam, method = "wss") # = 4

# Cluster
set.seed(43961)
pam.res <- cluster::pam(data_train, k = 2)

# Visual Clustering
factoextra::fviz_cluster(pam.res, data = data_train, geom = "point") + 
  labs(title = "Clustering Emissions from Gas Turbine\nwith K-Medoids/PAM (k = 2)")

# Print components of pam.res
print(pam.res)

# Cluster Validation
## Internal Validation
### Dunn's index
set.seed(43961)
clmethods <- c("pam")
intern <- clValid::clValid(data_train, nClust = 2:4, clMethods = clmethods, validation = "internal", maxitems = nrow(data_train))
summary(intern)

##              Score     Method    Clusters
## Connectivity	282.3984	pam	      2
## Dunn	        0.0130	  pam       2
## Silhouette	  0.3437	  pam       2
```

## 2. CLARA

```{r, message=FALSE, warning=FALSE}
# Choose optimum number of cluster
set.seed(43961)
factoextra::fviz_nbclust(data_train, FUNcluster = clara, method = "silhouette") #k = 2
factoextra::fviz_nbclust(data_train, FUNcluster = clara, method = "wss") #k = 5

# Compute CLARA
set.seed(43961)
clara.res <- cluster::clara(data_train, k = 2, metric = "euclidean", stand = TRUE)

# Print components of clara.res
print(clara.res)

# Visual Clustering
factoextra::fviz_cluster(clara.res, data_train, geom = "point") + 
  labs(title = "Clustering Emissions from Gas Turbine\nwith CLARA (k = 2)")

# Cluster Validation
## Internal Validation
### Dunn's index
set.seed(43961)
clmethods <- c("clara")
intern <- clValid::clValid(data_train, nClust = 2:5, clMethods = clmethods, validation = "internal", maxitems = nrow(data_train))
summary(intern)

##              Score     Method    Clusters
## Connectivity	345.8591	clara	      2
## Dunn	        0.0102	  clara       2
## Silhouette	  0.3407	  clara       2
```

# Comparison of clustering qualities among algorithms

```{r, message=FALSE, warning=FALSE}
set.seed(43961)
methodstocompare = c("pam", "clara")
intern <- clValid::clValid(data_train, nClust = 2:5, clMethods = methodstocompare,
   validation = "internal", metric = "euclidean", method = "average", maxitems = nrow(data_train))
clValid::optimalScores(intern)

##Best algo for clustering -> PAM / K-Medoids
##              Score         Method    Clusters
## Connectivity	282.39841270	pam	      2
## Dunn	        0.01304795	  pam       2
## Silhouette	  0.34366917	  pam       2

#Value for Connectivity width is quite high while Dunn index and Silhouette scores are low.
```

#Conclusion

Using the K-Medoids/PAM algorithm in clustering Emissions from Gas Turbine data is better than using the CLARA algorithm.
By comparison:

##              Score     Method    Clusters
## Connectivity	282.3984	pam	      2
## Dunn	        0.0130	  pam       2
## Silhouette	  0.3437	  pam       2

##              Score     Method    Clusters
## Connectivity	345.8591	clara	      2
## Dunn	        0.0102	  clara       2
## Silhouette	  0.3407	  clara       2

PAM Connectivity is lower than CLARA, that means PAM is better than CLARA
PAM Dunn Index is bigger than CLARA, that means PAM is better than CLARA
PAM Silhouette is closer to one than CLARA, that means PAM is better than CLARA

However, in terms of speed, CLARA clustering is faster because it is designed to cluster with large amounts of data.